---
title: "Lab 2"
author: "Phu Nguyen"
date: "9/2/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1

* The code I typed in the console:
```{r, eval=FALSE}
install.packages("ISLR2")
```
* Code chunks:
```{r}
library(ISLR2)
head(Boston)
```
* We are wanting to find a linear model with '*medv*' (median house value per $1000) as the response (output) and ‘*rm*' (average number of rooms per dwelling) as the predictor (input).  

* **Question 1**: For the 6th suburb of Boston what is the median house value and the average number of rooms per dwelling?  

  + **Answer**: For the 6th suburb of Boston the median house value '*medv*' is **28.7** and the average number of rooms per dwelling '*rm*' is **6.430**  
  
## Task 2

* The code I typed for the graph:
```{r, eval=TRUE, echo=TRUE}
plot(Boston$rm,Boston$medv,xlab = "rm",ylab = "medv")
```  

* **Question 2**: According to the plot what is the relationship between median value of homes and average number of rooms per dwelling?  

  + **Answer**: The relationship between '*medv*' and '*rm*' is *strong positive linear*. 
  + We can also find the correlation to explain the correlation between median value of homes and average number of rooms per dwelling by: 
```{r, eval=TRUE}
cor(Boston$rm,Boston$medv)
```  

## Task 3 

* In the code chunk type:
```{r, eval=TRUE}
lm.fit <- lm(medv ~ rm, data = Boston)
summary(lm.fit)
```

* **Question 1**: Give the linear model equation.  
  + **Answer**: $\hat{y} = 9.102x - 34.671$


* **Question 2**: What is the percent of variation of medv that can be explained by this model?  
  + **Answer**:  The percent of variation can be explain by R-square which is **0.4835**, therefore, **48.35%** of the data can be explained by the equation.  

  
* **Question 3**: Is rm a good predictor for medv? Justify your answer.  
  + **Answer**: Using $H_0: \beta_1 = 0$ and p-value = 0 
  + Since p-value < $\alpha = 0.05$, we fail to $RHo$ and therefore there is a relationship between *medv* and *rm*.  

## Task 4

* In a code chunk type:

```{r, eval=TRUE, echo=TRUE}
confint(lm.fit)
```   

* **Question 6**: What is the 95% confidence interval for the slope $\beta_1$ of this model?  
  + **Answer**: [\$8,278.85, \$9,925.36]  

## Task 5

* The *predict()* function can be used to produce predictions, confidence interval and prediction intervals for the prediction of *medv* for a given value of *rm*.
* The **confidence interval** is used to determine the average predicted value for the response variable.
* The **prediction interval** is used to determine the prediction for one observation of the response
variable.
* Suppose we want to determine a predicted value of *medv* based on the average number of rooms per dwelling at 5, 6, and 7. We can type the following in a code chunk  

```{r}
predict(lm.fit, data.frame(rm = c(5, 6, 7)))  
predict(lm.fit, data.frame(rm = c(5, 6, 7)),
        interval = "confidence")  
predict(lm.fit, data.frame(rm = c(5, 6, 7)),
        interval = "prediction")  
```  

* **Question 7**: What is the predicted median value of homes where the average number of rooms per dwelling is 5?  
  + **Answer**: \$10,839.92 

* Notice that the **confidence interval** for 5 is [9.634, 12.045]. The interpretation is: on average the median value of the homes in all of the suburbs with average of 5 rooms is between \$9,634 and \$12,45.
* Notice that the **prediction interval** for 5 is [-2.214, 23.894]. The interpretation is: if we look at one
suburb, the predicted median home value for that suburb will be between -\$2,214 and \$23,894.  

## Task 6

* We can check assumptions through the plots of the model.
* Using the code chunk type:  

```{r, eval=TRUE, echo=TRUE}
par(mfrow = c(2,2))
plot(lm.fit)
```  

* **Question 8**: Do there appear to be extreme values?
  + **Answer**: Yes at 366. Because we have extreme value, this might be the reason why our R-square value is low.  



* We can use the leverage statistics to determine extreme values. The function to fnd the leverage
statistics **hatvalues()**.
* Using the code chunk type:  

```{r}
which.max(hatvalues(lm.fit))
```  

* The **which.max()** function identifies the index (row) of the largest element of a vector.  


* **Question 9**: Which row has the largest leverage?  
  + **Answer**: 366  
  
* Using the code chunk type: **Boston[number of largest leverage,]**.  

```{r, eval= TRUE, echo=TRUE}
Boston[366,]
```  

* **Question 10**: How many average number of rooms per dwelling and what is the median value of the
homes in this suburb?  
  + **Answer**: The average number of rooms per dwelling '*rm*' is **3.561** and the median value of the homes '*medv*' is **27.5**.  
  




  









